---
title: "connectR Syntax"
author: "Dyfan Jones"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Quick Overview of connectR Syntax
### Connect to Databases Using `src_connectR`
```{r packages, results="hide", warning= FALSE, message= F}
library(tidyverse)
library(connectR)

src_connectR("PostgreSQL35W")->post
```

### Push R Dataframes into TeraData Using `copy_to`
```{r, fig.show='hold'}
mtcars->mtcars1
mtcars1$car<-rownames(mtcars1)
rownames(mtcars1)<-NULL

#Note. copy_to defaults temporary=FALSE
copy_to(post,mtcars1,"drj_mtcars",indexes = list(c("mpg","hp")),overwrite=T)
```

### Select Columns Using `select`
```{r, fig.show='hold', results =F}
drj_mtcars<-tbl(post,"drj_mtcars") %>% collect() 
drj_mtcars %>% select(car, contains('a'), ends_with('s')) %>% head
```
```{r, echo=F}
drj_mtcars %>% select(car, contains('a'), ends_with('s'))->drj_knitr1
knitr::kable(head(drj_knitr1))
```

### Extract Distinct Rows Using `distinct`
```{r, fig.show='hold'}
drj_mtcars %>% select(am) %>% distinct
```

### Add New Columns Using `mutate` 
```{r, fig.show='hold', results =F}
drj_mtcars %>% select(-mpg,-wt,hp) %>% 
  mutate(gear_no=ifelse(gear<=5,"high","low"),
         car_type=ifelse(grepl("merc",tolower(car)),"nice","ordinary"),
         cyl_sq=cyl^2,
         row_no=row_number(drat)) %>% head
```
```{r, echo=F}
drj_mtcars %>% select(-mpg,-wt,hp) %>% 
  mutate(gear_no=ifelse(gear<=5,"high","low"),
         car_type=ifelse(grepl("merc",tolower(car)),"nice","ordinary"),
         cyl_sq=cyl^2,
         row_no=row_number(drat))->drj_knitr2
knitr::kable(head(drj_knitr2))
```

### Order Data Using `arrange` 
```{r, fig.show='hold', results =F}
drj_mtcars %>% arrange(mpg)
drj_mtcars %>% arrange(cyl, desc(mpg))
```
```{r, echo=F}
knitr::kable(head(drj_mtcars %>% arrange(cyl, desc(mpg))))
```

### Aggregate Rows Using `summarise` 
```{r, fig.show='hold', results =F}
drj_mtcars %>% summarise(mpg_mean = mean(mpg),
                         max_disp = max(disp))
```
```{r, echo=F}
knitr::kable(drj_mtcars %>% summarise(mpg_mean = mean(mpg),
                         max_disp = max(disp)))
```

### Group Operations Using `group_by` 
```{r, fig.show='hold', results=F}
drj_mtcars %>% group_by(cyl) %>%
  summarise(mpg_mean = mean(mpg),
            max_disp = max(disp))
```
```{r,echo=F}
knitr::kable(drj_mtcars %>% group_by(cyl) %>%
  summarise(mpg_mean = mean(mpg),
            max_disp = max(disp)))
```
```{r, fig.show='hold', results=F}

drj_mtcars %>% group_by(cyl) %>% 
  mutate(row_number = row_number(car)) %>% 
  filter(row_number <= 3) %>% 
  select(cyl, row_number) %>% 
  arrange(cyl, row_number)
```
```{r,echo=F}
knitr::kable(drj_mtcars %>% group_by(cyl) %>% 
  mutate(row_number = row_number(car)) %>% 
  filter(row_number <= 3) %>% 
  select(cyl, row_number) %>% 
  arrange(cyl, row_number) %>% head)
```

### Drop Data from database using `db_drop_table`
```{r, fig.show='hold'}
db_drop_table(post$con,"drj_mtcars")
```

### Sending queries to the databse using `db_send_query`
```{r, fig.show='hold'}
db_send_query(post,"create table kaggle_test as
                        (Select
                        a.name,
                        a.style,
                        a.abv,
                        b.brewery_name,
                        b.city
                        from kaggle_beers a
                        inner join kaggle_breweries b
                        on a.brewery_id = b.brewery_id) with no data;")

#Tidying up the database
db_send_query(post,"drop table kaggle_test")
```

