---
title: "Lazy Evaluation Translator"
author: "Dyfan Jones"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Intro

connectR has a built in translator that will translate R code into SQL so a database can understand R_Code.

First lets connect to the database and check connectR's translator.

```{r, fig.show='hold'}
library(tidyverse, warn.conflicts = FALSE)
library(connectR)

src_connectR("PostgreSQL35W")->post
```

Lets check the default translator dplyr provides, and compared it to connectR's translator

```{r, fig.show='hold'}
# Default translator
sql_translate_env()

# connectR's translator
sql_translate_env(post$con)
```

As we can see connectR has a large tanslator that tries to cover as much as possible.

## Scalar functions

#### Basic Operators
 R_code | SQL Code 
--------|----------
       !|       NOT  
      !=|        <>                            
    %in%|        IN             
   &, &&|       AND             
 `|, ||`|        OR   
       ^|     POWER                     
      ==|         =   
      
#### Casting functions
 R_code | SQL Code | Description
--------|----------|------------------------------------------------
 as.bigint | BIGINT| `as.bigint(x)`: x is variable to be cast
 as.character | CHAR | `as.character(x,y)`: x is variable to be cast, y length of character. Returns `... x CHAR(y)`
 as.Date | DATE | `as.Date(x)`: x is variable to be cast
 as.decimal | DECIMAL | `as.decimal(x,pre,sca)`: x is variable to be cast, pre and sca are parameters to define Decimal dimensions. Returns `... x DECIMAL (pre,sca)`
 as.double | NUMERIC | `as.double(x)`: x is variable to be cast            
 as.float | FLOAT | `as.float(x)`: x is variable to be cast
 as.int | INT | `as.int(x)`: x is variable to be cast   
 as.integer | INT | `as.integer(x)`: x is variable to be cast   
 as.logical | BOOLEAN | `as.logical(x)`: x is variable to be cast       
 as.numeric | NUMERIC | `as.numeric(x)`: x is variable to be cast       
 as.smallint | SMALLINT | `as.smallint(x)`: x is variable to be cast   
 as.timestamp | TIMESTAMP | `as.timestamp(x)`: x is variable to be cast
 as.tinyint | TINYINT | `as.tinyint(x)`: x is variable to be cast
 as.varchar | VARCHAR | `as.varchar(x,y)`: x is variable to be cast, y length of varchar. Returns `... x VARCHAR(y)`          
 
#### Functions (that differ)
 R_code | SQL Code | Description
--------|----------|------------------------------------------------
 paste  | CONCAT   | `paste(x1,x2,...)`: Returns `CONCAT(x1,x2,...)`
 paste0 | `||`     | `paste0(x1,x2,...)`: Concatenates variables, strings, numbers, etc... together by using piping. Returns `x1||x2||...` (Preferred method due to issue in Teradata.)
 grepl  | LIKE     | `grepl(x,y)`: x pattern, y variable. Returns `... y LIKE x`
 trimws | TRIM     | `trimws(x)`: x variable to be trimmed.
 ifelse | CASE WHEN | `ifelse(x,...,...)`: Returns a case statement  `CASE WHEN x ... ELSE ... END`
 if_else | CASE WHEN | `ifelse(x,...,...)`: Returns a case statement  `CASE WHEN x ... ELSE ... END`
 if(){}else{} | CASE WHEN | `if(x){...}else{...}`: Returns a case statement with an else `CASE WHEN x ... ELSE ... END`
 tolower | LOWER   | `tolower(x)`: Returns `LOWER(x)`
 toupper | UPPER   | `toupper(x)`: Returns `UPPER(x)`
 scan    | STRTOK  | `scan(x,"-",1)`: Aligns to SAS scan function, returns `STRTOK(x,1,"-")`

#### Lubridate style functions
 R_code | SQL Code | Description
--------|----------|------------------------------------------------
 yeardiff |          | `yeardiff(x,y=NULL)`: Calculates the year difference between to date/timestamp fields. x initial date/timestamp, y latest date/timestamp (if y is null then y = DATE)
 monthdiff |          | `monthdiff(x,y=NULL)`: Calculates the month difference between to date/timestamp fields. x initial date/timestamp, y latest date/timestamp (if y is null then y = DATE)
  datediff |          | `datediff(x,y=NULL)`: Calculates the date difference between to date/timestamp fields. x initial date/timestamp, y latest date/timestamp (if y is null then y = DATE)
 d_wk   | TD_DAY_OF_WEEK | `d_wk(x)`: Returns day of the week as a numeric                       
 d_mth   | TD_DAY_OF_MONTH | `d_mth(x)`: Returns day of the month as a numeric
 wd_mth | TD_WEEKDAY_OF_MONTH | `wd_mth(x)`: Returns weekday of the year as numeric                    
 w_mth | TD_WEEK_OF_MONTH | `w_mth(x)`: Returns week of the month as numeric
 d_yr   | TD_DAY_OF_YEAR | `d_yr(x)`: Returns day of the year as a numeric
 w_yr  | TD_WEEK_OF_YEAR | `w_yr(x)`: Returns week of the year as a numeric
 m_yr       | TD_MONTH_OF_YEAR | `m_yr(x)`: Returns month of the year as a numeric
 year   | TD_YEAR_OF_CALENDAR | `year(x)`: Returns year as numeric

**NOTE: To check all functions that can be translated please check `sql_translate_env(zeus$con)`**
